# Wastewater treatment plants contain a conserved core community of bacteria

```{r "setup", warning=FALSE}
source("R/functions.R")
source("R/core_community.R")
sessionInfo()
```

# Data

The source dataset contains all:

  1. Q3 samples from 2008 and 2008 from all plants. 
  2. time series from AAW
  
```{r "Load data", warning=FALSE}
identities      <- as.character(c(94, 97, 99))
fname_template  <- "data/otuXX/seqs_XX_otutable.biom"
filenames       <- sapply(identities, 
                          function(id) gsub("XX", id, fname_template) ) 

datasets        <- lapply(identities, function(identity) 
                               LoadData(biompath = filenames[identity], 
                                        mapfpath = "data/mapfile.txt") )
names(datasets) <- paste0("otu", identities)

print(datasets_summary <- data.frame(
  "samples" = sapply(datasets, function(identity) sum(nsamples(identity))),
  "reads"   = sapply(datasets, function(identity) sum(sample_sums(identity))),
  "OTUs"    = sapply(datasets, function(identity) sum(ntaxa(identity))) 
     ))
```

### Datasets

 1) `coreDatasets`  Two samples from each plant from the summer 2008 and 2009 were used to calculate the core microbial community in the cross-section of Danish plants. 
 2) `tsDatasets`  All the samples from Aalborg West from 2006 and 2010 were used to calculate the core community in the time-series.

```{r "Subset datasets", warning=FALSE}
subsample_depth <- 40000
# List containing phyloseq objects for core dataset at 94, 97 and 99% identity 
coreDatasets    <- lapply(datasets, 
                          function(identity) selectCoreDataset(identity, 
                                                      depth = subsample_depth, 
                                                      seed  = 1234 ))
names(coreDatasets) <- paste0("otu", identities)
lapply(coreDatasets, function(identity) printDatasetStats(identity) )

# List containing phyloseq objects for time-series dataset at 94, 97 and 99% id 
tseriesDatasets <- lapply(datasets, 
                  function(identity) selectAAWDataset(identity, 
                                                      depth = subsample_depth))
names(tseriesDatasets) <- paste0("otu", identities)
lapply(tseriesDatasets, function(identity) printDatasetStats(identity) )
```

## 1. Core community

## Quantifying the core community

```{r "Figure 1: Core OTU conservation plot", warning=FALSE, message=FALSE}
coredataframe  <- calcSummaryData(coreDatasets, identities, core_cutoff=1)

coredata_by_id <- group_by(coredataframe, identities)
summarise(coredata_by_id, 
          coretotalOTUs    = sum(taxsum),
          coreOTUs         = sum(taxsum[corestatus == "core"]),
          percentcorereads = round(readprop[corestatus == "core"] * 100, 1))

coreplots <- plotCore(coredataframe)
otuplot   <- coreplots[[1]]
readsplot <- coreplots[[2]]
print(Figure1   <- grid.arrange(otuplot, readsplot, nrow = 2))

pdf(file = "figs/Figure_1_core_otu_conservation40k.pdf")
Figure1
dev.off()
```


### Fraction of Frequently occuring OTUs (observed 26 times)

```{r "Count frequently observed OTUs (nObs == 26 )"}
filter(coredataframe, numsamples == 26 ) %>%
  group_by(identities) %>%
  summarise(
    total_OTUs  = sum(taxsum),
    percent_reads = round(sum(readprop) * 100, 1) )
```

### Fraction of Frequently occuring OTUs (observed 20 + times)

```{r "Count frequently observed OTUs (nObs > 20 )"}
filter(coredataframe, numsamples >= 20 ) %>%
  group_by(identities) %>%
  summarise(
    total_OTUs  = sum(taxsum),
    percent_reads = round(sum(readprop) * 100, 1) )
```

### Fraction of Infrequently observed OTUs (observed no more than 5 times)

```{r "Count infrequently observed"}
    filter(coredataframe, numsamples < 6 ) %>%
  group_by(identities) %>%
  summarise(
    total_OTUs  = sum(taxsum),
    percent_reads = round(sum(readprop) * 100, 1) )
```


## Cumulative abundance distribution

The distribution of abundances in bacterial communities is uneven - some 
organisms are abundant some are rare.

Plotting percent cumulative abundance vs. rank ordered OTUs describes this 
distribution.

```{r "Figure 2: Cumulative abundance curve", warning=FALSE, message=FALSE}
cum_abun_plot <- plot_CumulRankAbundance(dataset = coreDatasets[["otu94"]], 
                                       fname = "figs/figure2_cum_abun_plot_94.pdf",
                                       dominant_fraction = 0.8)
print(cum_abun_plot)
```

### Distribution of abundances per OTU

There is a lot of variation in the abundances but many of the abundant organisms are consitently abundant. These are the organisms that are core. The size of the core will increase as a function of depth for these consistent organisms.


```{r "Figure 3: boxplot the top OTUs"}
plotTopN(coreDatasets[["otu94"]], 
         topN = 50, 
         plot_filename = "figs/Figure_3_top50_id94_boxplot.pdf", 
         core_cutoff = 1.0)
```

### Temporal stability of a single plant through time

Time series in AAW

What proportion of the core OTUs are core in: 
 - the two AAW samples used for the core
 - all the AAW samples

```{r "calc AAW timeseries core"}
tsdataframe <- calcSummaryData(tseriesDatasets, identities, core_cutoff=1)
group_by(tsdataframe, identities) %.%
  summarise( coretotalOTUs    = sum(taxsum),
             coreOTUs         = sum(taxsum[corestatus == "core"]),
             percentcorereads = round(readprop[corestatus == "core"] * 100, 1))

tscoreplot  <- plotCore(tsdataframe)
tsotuplot   <- tscoreplot[[1]]
tsreadsplot <- tscoreplot[[2]]
print(tscoreplot  <- grid.arrange(tsotuplot, tsreadsplot, nrow = 2))

pdf(file = "figs/Figure_S2_aawtimeseries_core_conservation_40k.png")
tscoreplot
dev.off()
```


### Binning of OTUs by observation frequency and frequency of high-abundance

```{r "Dominant OTUs"}
dataset   <- coreDatasets[["otu94"]] 
core_prop <- 1

dominant  <- sapply(sample_names(dataset), 
       function(samplename) fill_ha_data(dataset, samplename))
row.names(dominant) <- taxa_names(dataset)

observed <- as.data.frame(otu_table(transform_sample_counts(
    dataset, function(x) ifelse(x > 0, 1, 0 ) )))

summary.df <- data.frame(
                      "OTU"    = taxa_names(dataset),
                      "nHA"    = apply(dominant, 1, sum),
                      "nObs"   = apply(observed,      1, sum),
                      "median" = apply(otu_table(dataset), 1, median),
                      "geomean"= apply(otu_table(dataset), 1, function (x) 
                                          round(exp(mean(log(x))), 1)),
                      "max"    = apply(otu_table(dataset), 1, max),
                      "min"    = apply(otu_table(dataset), 1, min),
                      "n1per"  = apply(otu_table(dataset), 1, 
                                           function(x) sum(x > 400)) )

# how does median relate to nHA and nObs?
ggplot(summary.df, aes(nObs, median)) + 
  geom_point(alpha = 0.2) + 
  scale_y_log10()
ggplot(summary.df, aes(nHA, median)) + 
  geom_point() + scale_y_log10()
```

These plots are the justification for having nHA > 10 as the cutoff for significance. 

## Figure 4

Sets up the empty Figure 4 plot. The data was filled manually using Inkscape.

```{r "Figure 4: Dominant vs. Frequency"}
summary.df <- mutate(summary.df, 
                  group    = factor(cut(nHA, breaks=c(-1, 0, 9, 25, 26), 
                                  labels= c("4","3", "2", "1")), levels = 1:4),
                  Obsclass = cut(nObs, breaks=c(0, 19, 25, 26), 
                                  labels= c("ob1", "ob20", "ob26")) )
summary.df <- cbind(summary.df, as.data.frame(otu_table(dataset)))
summary.df <- arrange(summary.df, desc(nHA), desc(nObs), desc(median))

print(Figure4 <- plotNewFigure4(summary.df))

group_by(summary.df, group) %.%
  summarise( nOTUs = n() ,
             percent = round(sum(percent), 0)) 
group_by(summary.df, Obsclass) %.%
  summarise( nOTUs = n() ,
             percent = round(sum(percent), 0)) 

summary.df2 <- group_by(summary.df, nHA, nObs) %.%
  summarise( nOTUs = n() ,
             percent = round(sum(percent), 2)) 
summary.df2 <- mutate(summary.df2, 
                  group    = factor(cut(nHA, breaks=c(-1, 0, 9, 25, 26), 
                                  labels= c("4","3", "2", "1")), levels = 1:4),
                  Obsclass = cut(nObs, breaks=c(0, 19, 25, 26), 
                                  labels= c("ob1", "ob20", "ob26")) )

```

## Data for plotting onto Figure 4
```{r}
nObs_breaks = c(1, 20, 25, 26)
nHA_breaks  = c(0, 1, 10, 25, 26) 
library(scales)

grouprecs <- data.frame(
  group  = c(   4,    3,    2,    1),
  ystart = c(-0.5,  0.5,  9.5, 25.5),
  ystop  = c( 0.5,  9.5, 25.5, 26.5),
  xstart = c(-0.5, -0.5, 19.5, 25.5),
  xstop  = c(26.5, 26.5, 26.5, 26.5) )

grouprecs$group = factor(grouprecs$group)

A <- ggplot(data=summary.df, aes(y = nHA, x = nObs) ) +
  xlab("Observed in n samples") +
  ylab("Highly-abundant in n samples") +
  scale_fill_manual(values = alpha(c("red", "orange", "darkgreen", "grey"), 0.3)) +
  geom_rect( aes(x = NULL, y = NULL, xmin= xstart, xmax=xstop, 
                 ymin=ystart, ymax=ystop, fill = group), data=grouprecs) +
  geom_point(data=summary.df, stat= "identity", size = 2, fill = "black", alpha = 0.1) +
  scale_y_continuous(breaks= nHA_breaks,  labels= c(0, 1, 10, 25, 26), expand=c(0,0)) +
  scale_x_continuous(breaks= nObs_breaks, labels= c(1, 20, 25, 26), expand=c(0,0)) +
  xlab("Observed in n samples") +
  ylab("Highly-abundant in n samples") +
  theme_bw() +
  theme(legend.position = "none",
        axis.title.x = element_text(size = 6),
          axis.title.y = element_text(size = 6, vjust = 0.2),
          axis.text.x  = element_text(size = 5),
          axis.text.y  = element_text(size = 5),
          axis.ticks.x = element_line(size = 0.3),
          axis.ticks.y = element_line(size = 0.3),
          panel.grid.minor = element_blank(),
          panel.border     = element_rect(color= "black"),
          axis.line        = element_line(size = 0.3))

ggsave(filename="figs/Figure4.pdf", plot= A, width= 8, height = 6, units= "cm")

A2 <- ggplot(data=summary.df, aes(y = nHA, x = nObs) ) +
  xlab("Observed in n samples") +
  ylab("Highly-abundant in n samples") +
  scale_fill_manual(values = alpha(c("red", "orange", "darkgreen", "grey"), 0.3)) +
  geom_rect( aes(x = NULL, y = NULL, xmin= xstart, xmax=xstop, 
                 ymin=ystart, ymax=ystop, fill = group), data=grouprecs) +
  geom_point(data=summary.df, stat= "identity", size =(3), fill = "black", alpha = 0.2) +
  scale_y_continuous(breaks= nHA_breaks,  labels= c(0, 1, 10, 25, 26), expand=c(0,0)) +
  scale_x_continuous(breaks= nObs_breaks, labels= c(1, 20, 25, 26), expand=c(0,0)) +
  xlab("Observed in n samples") +
  ylab("Highly-abundant in n samples") +
  theme_bw() +
  theme(legend.position = "none", 
        axis.title.x = element_blank(),
        axis.text.x  = element_blank(),
        axis.title.y = element_text(size = 6, vjust = 0.2),
        axis.text.y  = element_text(size = 6),
        axis.ticks.x = element_line(size = 0.3),
        axis.ticks.y = element_line(size = 0.3),
        panel.grid.minor = element_blank(),
        panel.border     = element_rect(color= "black"),
        axis.line        = element_line(size = 0.3))

df3 <- group_by(summary.df, nHA) %.%
  summarise( nOTUs = n() ,
             percent = round(sum(percent), 0)) 

B <- ggplot(df3, aes(y = nHA, x = cumsum(percent)) ) +
  geom_line() +
  scale_y_continuous(breaks= nHA_breaks,  labels= c(0, 1, 10, 25, 26)) +
  xlab("Cumulative percentage") +
  ylab("Highly-abundant in n samples") +
  theme_bw() + 
  theme(axis.title.y = element_blank(),
         axis.text.y  = element_blank(),
        axis.title.x = element_text(size = 6),
          axis.text.x  = element_text(size = 6),
          axis.ticks.x = element_line(size = 0.3),
          axis.ticks.y = element_line(size = 0.3),
          panel.grid.minor = element_blank(),
          panel.border     = element_rect(color= "black"),
          axis.line        = element_line(size = 0.3))

df4 <- group_by(summary.df, nObs) %.%
  summarise( nOTUs = n() ,
             percent = round(sum(percent), 0)) 

C <- ggplot(df4, aes(y = cumsum(percent), x = nObs) ) +
  geom_line() +
  scale_x_continuous(breaks= nObs_breaks, labels= c(1, 20, 25, 26)) +
  xlab("Observed in n samples") +
  ylab("Cumulative percentage") +
  theme_bw() +
  theme(axis.title.x = element_text(size = 6),
          axis.title.y = element_text(size = 6, vjust = 0.2),
          axis.text.x  = element_text(size = 6),
          axis.text.y  = element_text(size = 6),
          axis.ticks.x = element_line(size = 0.3),
          axis.ticks.y = element_line(size = 0.3),
          panel.grid.minor = element_blank(),
          panel.border     = element_rect(color= "black"),
          axis.line        = element_line(size = 0.3))

plots <- list(A2, B, C)

layout <- matrix(c(1, 1, 1, 2, 1, 1, 1, 2, 3, 3, 3, 4), nrow = 3, byrow = TRUE)
multiplot(plotlist = plots, layout = layout)

pdf(file="figs/Figure4_multi.pdf", width= 6,2, height = 6.2 )
multiplot(plotlist = plots, layout = layout)
dev.off()


```



How many OTUs/reads are in each category

```{r "Figure 4 results for plot"}
r <- melt(select(summary.df, OTU, group, Obsclass, AMPA057:AMPA724),
                id.vars=c("OTU", "group", "Obsclass") , 
                variable.name="sample", value.name="count" ) %.%
  group_by(Obsclass, group) %.%
  summarise(readpercent = round((sum(count) / (26 * 40000)) * 100, 1),
                  nOTUs = n_distinct(OTU))
acast(r, group ~ Obsclass, value.var= "nOTUs", 
                           fun.aggregate= sum, 
                           margins= TRUE)
acast(r, group ~ Obsclass, value.var= "readpercent", 
                           fun.aggregate= sum, 
                           margins= TRUE)
```


```{r}
Acinetobacter <- subset_taxa(physeq = coreDatasets[["otu94"]], 
                                 Genus == "Acinetobacter")
ntaxa(Acinetobacter)
round(sum(taxa_sums(Acinetobacter)) / sum(taxa_sums(coreDatasets[["otu94"]])) * 100, 
      digits = 2)
```



```{r}
# NB rounding error# 3. transiently highly-abundant
group3otus <- as.vector(with(summary.df, summary.df[ group == 3, "OTU"]))
tHA_percent <- round(
  as.data.frame(otu_table(dataset)[ group3otus, ] / 40000 * 100), 1)


df       <- data.frame("rank" = 1:26, 
                       "percentHA" = sort(colSums(tHA_percent), decreasing=TRUE))
plant    <- samData(dataset)[row.names(df), "plant"]
df$plant <- as.character(plant$plant)  # crazy phyloseq object!!

#TODO fix the label on this? why does the sample_data object persist after as.character()
# ggplot(df, aes(x=rank, y= percentHA)) + geom_point(stat = "identity") + 
#   scale_x_discrete(labels= plant)
# + xlab("samples") + ylab("transiently HA read percent")

# num trans.abun OTUs per sample
apply(tHA_percent, 2, function(sample) sum(sample > 1)) 
samData(dataset)[names(sort(colSums(tHA_percent), decreasing=TRUE)), "plant"]

HA.tran <- subset(summary.df, nObs  <= 20 & nHA > 0 &  (n1per > 0))
tax_table(dataset) [ as.vector(HA.tran$OTU), 1:6 ]
nrow(tax_table(dataset) [ as.vector(HA.tran$OTU), 1:6 ])
```


## Single sequence resolution on the abundant OTUs

```{r}
# Checked OTU681 using parsimony insertion in ARB
otu94.per <- transform_sample_counts(coreDatasets[["otu94"]], 
                                     function(x) x / sum(x) * 100)
tax_table(otu94.per)["681", 5:6] <- c("Intrasporangiaceae", "Tetrasphaera_etal")
Tetrasphaera <- subset_taxa(otu94.per, Genus ==  "Tetrasphaera_etal" )
write.table(taxa_names(Tetrasphaera), file="data/unfiltered_tetra_otulist94.txt",
            quote= FALSE, col.names= FALSE, row.names= FALSE, sep="\n")


plot_heatmap(Tetrasphaera, sample.label="plant", method= "NMDS")
plot_bar(Tetrasphaera, "Genus", facet_grid=(year ~ plant), fill= "OTU")
```

Then the reads that make up these three OTUs were parsed out of the seqs file,
and reclustered at 99% identity using qiime.




# export the OTU data to a table for the Table S2

```{r "Export ecocore Table S2"}
#list of tax names 
eco.core.taxa     <- as.vector(with(summary.df, 
                               summary.df[ HAclass %in% c("HA10", "HA26"), "OTU"]))
trans.abun.noncore.taxa <- as.vector(with(summary.df, 
                               summary.df[ HAclass == "HA1" & n1per > 0, "OTU"]))

length(trans.abun.noncore.taxa)

# lists of taxnames ecocore and abun.noncore
taxa.to.export <- c(ecocore.taxa, trans.abun.noncore.taxa)
outstats       <- select(summary.df, 1:12) %.%
                  filter(OTU %in% taxa.to.export)
outtaxtable    <- as.data.frame(tax_table(coreDatasets[["data.94"]])[ taxa.to.export, 1:6])
outdata        <- cbind(outstats, outtaxtable) %.%
                  arrange(desc(median))

table(outdata$group)

write.table(outdata, file="figs/Table_S2_newcore_otutable.txt", sep="\t", 
            quote=FALSE, row.names=FALSE, col.names=TRUE)
```


### Nitrotoga and Nitrospira

```{r "NOB"}
# Compare replicate data

data.97 <- datasets[["otu97"]]
amplibs <- sample_data(data.97)$SampleID[
                      sample_data(datasets[["otu97"]])$sample_id == "293" ]
data.97 <- prune_samples(x=data.97, samples=as.character(amplibs) )
table(sam_data(data.97)$sample_id, sam_data(data.97)$dna_id)
data.97 <- rarefy_even_depth(data.97, 
                             rngseed = 1234,
                             sample.size = 14000, 
                             trimOTUs= TRUE)

data.97 <- transform_sample_counts(physeq = data.97, function(x) x / sum(x) * 100)
NOB <- subset_taxa(data.97, (Family ==  "Gallionellaceae") | 
                            (Genus == "Nitrospira") )

plot_bar(NOB, "Genus", facet_grid = . ~ SampleID  ) +
    geom_bar(aes(fill = Genus), stat = "identity", position = "stack")

plot_bar(NOB, "Genus", facet_grid = dna_id ~SampleID  ) +
    geom_bar(aes(fill = Genus), stat = "identity", position = "stack")

# compare core samples
data.97 <- coreDatasets[["otu97"]]
data.97 <- transform_sample_counts(physeq = data.97, 
                                   function(x) x / sum(x) * 100)
NOB <- subset_taxa(data.97, (Genus ==  "Nitrotoga_etal") | 
                            (Genus == "Nitrospira") )

p <- plot_bar(NOB, "Genus", facet_grid = year ~ plant ) +
    geom_bar(aes(fill = Genus), stat = "identity", position = "stack") 
levels(p$data$Genus) <- c("Nitrospira", "Nitrotoga")
p$data$plant_name <- factor(gsub(x=p$data$plant_name, 
                                 pattern="oe", replacement="ø"))
p$data$plant_name <- factor(gsub(x=p$data$plant_name, 
                                 pattern="aa", replacement="å"))

totals <- as.data.frame(
  select(p$data, OTU, Sample, Abundance, year, Genus, plant_name) %.%
  filter(year == 2009, Genus == "Nitrotoga") %.%
  acast(plant_name ~ Genus, value.var="Abundance", sum))
totals$plant_name    <- row.names(totals)
plant_names_by <- with(totals, 
              totals[ order(Nitrotoga, decreasing=TRUE), "plant_name"])
p$data$plant_name <- factor(p$data$plant_name, levels=plant_names_by)

# formatted in greyscale for publication
p2 <- ggplot(p$data, aes(x = plant_name, y = Abundance, fill = Genus)) + 
    geom_bar(position=position_dodge(), stat = "identity") +
    facet_grid(year ~ .) +
    theme_bw() +
    ylab(label= "Read abundance (%)") +
    xlab(label = "Plant") +
    scale_fill_manual(values=c("grey50", "black")) +
    annotate("text", x= "Aalborg East", y= 2.5, 
             label= "2008", size = 2) +
    theme(axis.title.x = element_text(size = 8),
          axis.title.y = element_text(size = 8, vjust = 0.2),
          axis.text.x  = element_text(size = 6,  hjust = 1, vjust = 1, 
                                      angle = 45),
          axis.text.y  = element_text(size = 6),
          axis.ticks.x = element_line(size = 0.3),
          axis.ticks.y = element_line(size = 0.3),
          panel.grid.minor = element_blank(),
          strip.background = element_rect(linetype= "blank", fill = "white"),
          strip.text.y     = element_blank(),
          axis.line        = element_line(size = 0.3),
          legend.title     = element_blank(),
          legend.text      = element_text(size=5, face="italic"),
          legend.key.size  = unit(0.4, "lines"),
          legend.key       = element_rect(size= 1, colour= "white"),
          legend.justification = c(1, 1), 
          legend.position      = c(1.055, 1.086)
          )

fname <- "figs/Figure_7_NOB_core.pdf"
ggsave(plot=p2, file = fname, width=8, height=6.37, units="cm")

# Compare time series data
ts.97 <- tseriesDatasets[["otu97"]]
ts.97 <- transform_sample_counts(physeq = ts.97, function(x) x / sum(x) * 100 )
NOB <- subset_taxa(ts.97, (Genus ==  "Nitrotoga_etal") | (Genus == "Nitrospira") )
NOB <- tax_glom(NOB, taxrank = "Genus")


p <- plot_bar(NOB, "date") +
  geom_bar(aes(fill = Genus), stat = "identity", position = "stack")
levels(p$data$Genus) <- c("Nitrospira", "Nitrotoga")

p$data$month <- factor(p$data$month, labels= c("February", "May", "August"))
p$data$date  <- as.Date(p$data$date, format = "%d-%m-%Y")


p2 <- ggplot(p$data, aes(x = date, y = Abundance, color = Genus)) + 
  geom_point( stat = "identity") +
  scale_x_date( ) +
  scale_y_continuous(limit = c(0, 2.1)) +
  geom_line() +
  scale_color_brewer(palette = "Set1") +
  theme_bw() +
  ylab(label= "Read Abundance (%)") +
  xlab(label = "Plant") +
  theme(axis.title.x = element_text(size = 18),
        axis.title.y = element_text(size = 18, vjust = 0.2),
        axis.text.x  = element_text(size = 16,  hjust = 1, vjust = 1, 
                                    angle = 45),
        axis.text.y  = element_text(size = 16),
        axis.ticks.x = element_line(size = 0.3),
        axis.ticks.y = element_line(size = 0.3),
        strip.text   = element_text(size = 16),
        panel.grid.minor = element_blank(),
        strip.background = element_rect(linetype= "blank", fill = "white"),
        axis.line        = element_line(size = 0.3),
        legend.title     = element_blank(),
        legend.text      = element_text(size=18, face = "italic") )

ggsave(plot=p2, path = "figs/", file = "NOB_timeseries.pdf", width=20, height = 12, units="cm")


d <- read.delim('data/MIDAS_combined.csv', row.names = 1)
mean(d$T_lc, na.rm= TRUE)

# TODO change data to Table S1
# yearly temp means in DK plants
d.means <- ddply(d, .(Plant, Quarter), summarize, 
              mean = mean(T_lc, na.rm= TRUE),
              sd   = sd(T_lc, na.rm= TRUE),
              n    = sum(!is.na(T_lc) ))
d.means <- d.means[ d.means$n > 3 , ]
d.means <- subset(d.means, Quarter %in% c(1, 3) )

ggplot(d.means, aes(Plant, mean, color = factor(Quarter))) + geom_point() +
  theme(axis.text.x = theme_text(angle = 45, hjust = 1))

summarySE(d.means, measurevar="mean", groupvars=c("Quarter"), na.rm=TRUE)
```


```{r}
# what percentage of reads have a genus level classification?
data.97 <- coreDatasets[["otu97"]]
data.97.genus <- tax_glom(data.97, taxrank= "Genus", NArm= FALSE)
ntaxa(data.97.genus)
total_reads <- sum(sample_sums(data.97.genus))
data.97.genus_na <- subset_taxa(data.97.genus, is.na(Genus) )
ntaxa(data.97.genus_na)
total_genusna_reads <- sum(sample_sums(data.97.genus_na))
total_genusna_reads / total_reads

```
