# 2. Immigration and netgrowth
```{r}
source('R/functions.R')
source("R/core_community.R")
source("R/net_growth_rate.R")
sessionInfo()
```

## load data

```{r "Load data"}
biomfname = "data/otu97/netgrowth_seqs_97_otutable.biom"
netgrowthdata <- LoadData(biompath=biomfname, 
                          mapfpath="data/mapfile.txt")
```


```{r "wastewater core"}
ids    <- as.character(c(94, 97, 99))
fname_template <- "data/otuXX/netgrowth_seqs_XX_otutable.biom"
fnames <- sapply(ids, function(id) gsub("XX", id, fname_template) ) 

datasets <- lapply(ids, function(id) 
  LoadData(biompath= fnames[id], mapfpath= "data/mapfile.txt") )
ids <- sapply(ids, function(id) paste0("otu", id ))
names(datasets) <- ids

sapply(datasets, function(dataset)     sample_sums(dataset) )
sapply(datasets, function(dataset) sum(sample_sums(dataset)))
sapply(datasets, function(dataset) sum(   nsamples(dataset)))
sapply(datasets, function(dataset) sum(      ntaxa(dataset)))

min(sapply(datasets, function(dataset) sample_sums(dataset) ))
resample_depth <- 40000

rarefyDataset <- function(mydata, depth= FALSE, seed= FALSE){
  mydata <- subset_samples(mydata, type == "WW")
  mydata <- rarefy_even_depth(mydata,     rngseed = seed,
                                      sample.size = depth, 
                                         trimOTUs = TRUE)
  mydata
}

WWDatasets <- lapply(ids, function(id) 
                    rarefyDataset(datasets[[id]],
                                  depth = resample_depth, 
                                   seed = 1234))
names(WWDatasets) <- ids

lapply(WWDatasets, function(dataset) printDatasetStats(dataset) )

data.k <- SampleEvenDepth(datasets[["otu97"]], 
                          depth = 20000, 
                          seed = 1234)
print(sample_sums(data.k))
```

## Core community in influent wastewater

```{r "influent core community"}
coredataframe <- calcSummaryData(WWDatasets, ids, core_cutoff=1)

OTUtotals <- group_by(coredataframe, identities) %.%
  summarize( "totalOTUs"  = sum(taxsum),
     "coreOTUs"   = taxsum[ corestatus == "core"] , 
     "percentcoreOTUs"= round(taxsum[corestatus == "core"] / 
                               sum(taxsum) * 100, 1),
     "percentreads"   = round(readprop[corestatus == "core"] *
                                100, 1))
print(OTUtotals)

p <- plotCore(coredataframe, cumulative = FALSE)
otuplot   <- p[[1]]
readsplot <- p[[2]]

pdf(file = "figs/ww_core40k.pdf")
grid.arrange(otuplot, readsplot, nrow = 2)
dev.off()
grid.arrange(otuplot, readsplot, nrow = 2)
```


## Abundnace of top 10 OTUs

```{r "abundance of top10 in infleunt and sludge"}
ds   <- WWDatasets[["otu94"]]
ds.p <- transform_sample_counts(ds, function(x) x / sum(x))
n    <- 10
topNww_otus  <- names(sort(taxa_sums(ds.p), decreasing=TRUE)[1:n])
topN.ww      <- subset_taxa(ds, taxa_names(ds) %in% topNww_otus )

print(formatAverage(sample_sums(topN.ww) / sample_sums(ds) * 100,
                    decimals=0))

plot3 <- compareTop20AsWw(data.k)
ggsave(path="figs", filename="top20OTUsWW_vs_AS.pdf", 
       plot=plot3, width=8)
plot3
```



## Overall comparison of the diversity in the wastewater and activated sludge samples.

```{r "ordination"}
ord.cca <- ordinate(physeq = data.k, 
                    method = "CCA", 
                    distance = "bray")
print(
  plot1  <- plot_ordination(physeq = data.k, 
                            ordination=ord.cca, 
                            shape = "plant",
                            type = "samples", 
                            color = "sample_type") +
  geom_text(aes(label = pair), color= "black", size = 5, vjust = 2) +  
  geom_point(size = 5)
)

ggsave(path="figs", filename="CCA_AS_vs_WW.pdf", plot=plot1)
```


The CCA splits the samples by sample type in the first CA and roughly by plant in the second.


Add run model

## Calculate Net growth rates

```{r "Run Net growth rate script"}

df.netgrowth <- calcNetGrowthData(data.k, k_lower_bound=-0.2, seed = 1234)
df.netgrowth$km_class <- cut(df.netgrowth$km, 
                              breaks=c(-0.22, -0.1, 0, 0.03, 10), 
                              labels=c( "inactive", "low", "active", "max") )
df.netgrowth[(df.netgrowth$prop_ww == 0), "km_class" ] <- "max"

df.netgrowth$tax_name <- makeTaxLabels(df.netgrowth$OTU, data.k)

df.netgrowth$taxlab   <- ifelse((df.netgrowth$prop_ww > 0.01) &
                                 (df.netgrowth$km > -0.2) &
                                 (df.netgrowth$km <  0.01), 
                                  as.character(df.netgrowth$tax_name), "")

plist        <- plotNetGrowthDistribution(df.netgrowth)

print(plist[[1]])
print(plist[[2]]) 

ggsave(path="figs", filename="cumabun_by_netgrowth.pdf", 
       plot=plist[[2]], width=8, height=6, units="cm")


```

## Plot simple ratio 

```{r "Plot simple ratio"}
ggplot(data=df.netgrowth, aes(x = ratio, fill = class)) +
    geom_histogram() +
    labs(x = "ratio", y = 'OTU count') +
    theme(legend.position="none")
  
df.netgrowth  <- arrange(df.netgrowth, ratio)
  
ggplot(data=df.netgrowth, aes(x = ratio, y = cumsum(prop_as / 6) * 100)) +
    geom_step() +
    xlab("ratio") +
    ylab('Cumulative read abundance (%)')
```

#Summarise data per plant by class

Classes are "inactive", "low", "active", "max"

```{r Summarize % and n by class}

a <- group_by(df.netgrowth, pair, km_class) %.%
     summarise( AS_percent = sum(prop_as * 100),
                n_otus     = n() )

# mean sludge% by class
dcast(a, pair ~ km_class, value.var= "AS_percent", 
      fun.aggregate= function(x) round(mean(x), 1))
# mean # OTUs by class
dcast(a, pair ~ km_class, value.var= "n_otus", fun.aggregate= sum, margins= TRUE )

# percent reads near k max
cumsum(df.netgrowth$prop_as / 6)[sum(df.netgrowth$k > 0.02857 )]
high_growth_rate <- df.netgrowth[df.netgrowth$k > 0.02857, ]
max(high_growth_rate$prop_ww)

# percent of reads with k > -0.15
cumsum(df.netgrowth$prop_as / 6)[sum(df.netgrowth$km < -0)]

```

The assumption is that the net growth rate at steady state will be equal to the 
max net growth rate (1/35 = 0.0286).

There are a lot of OTUs that are at or near the upper limit. 

```{r}
printNetGrowthStats(df.netgrowth)
```

so the are on average 64% of the reads in the sludge detected in the influent, 
but this number is highly variable probably because the sampling of the sludge
taxa in the nfluent is near the level of detection.

```{r "color by influent abundance"}
myBreaks <- c(1,10)

ggplot(data=df.netgrowth, aes(x=prop_as*100, y= km, 
                               color = prop_ww*100, label = taxlab)) +
  geom_point() +
  geom_text( size = 3, color = "black", vjust= 1, hjust = 1) +
  scale_x_log10(breaks= c(0.01,0.1,1,10), limits= c(0.01,15)) +
  scale_colour_gradient(trans= "log", low="white", high="red",
                        breaks= myBreaks, labels = myBreaks) +
  facet_wrap(pair~plant) +
  labs(y= expression(paste("Bounded net growth rate (", d^-1, ")", sep="")), 
       x= 'biomass abundance (%)', 
       colour = "ww abundance (%)") 
ggsave(filename="figs/sludge_vs_k_all.png")


myBreaks <- c(1,10)

df.netgrowth$taxlab <- NA
df.netgrowth$taxlab <- with(df.netgrowth, 
                             ifelse((prop_ww >= 0.01) & (prop_as >= 0.0012) , 
                                    tax_name, taxlab) )

p <- plot_netgrowth(mydata=df.netgrowth, value="AAE-2")
p

info <- c(as.character(sample_data(data.k)$pair), "all")
per_plant_plots <- lapply(info, FUN=plot_netgrowth)
names(per_plant_plots) <- info
```

```{r "Percentage of AS reads also in influent"}
class_by_pair <- group_by(df.netgrowth, pair, class)
d <-summarise(class_by_pair, "total" = n()) %.%
    arrange(class, pair)
d1 <- d[d$class == "Both", "total"] / d[d$class == "AS", "total"] * 100
names(d1) <- unique(d$pair)
# result per pair
d1

by_class <- group_by(df.netgrowth, class)
d2 <-summarise(by_class, "total" = n()) %.%
    arrange(class)
d2
# overall fraction
d2[d2$class == "Both", "total"] / d2[d2$class == "AS", "total"] *100

```


## NOB in influent and effluent

```{r}
data.p <- transform_sample_counts(physeq = data.k, function(x) x / sum(x) * 100)
NOB <- subset_taxa(data.p,
                   (Genus ==  "Nitrotoga_etal") | (Genus == "Nitrospira") )
NOB.prop <- transform_sample_counts(physeq = NOB, function(x) x / sum(x) * 100 )

p <- plot_bar(NOB, "Genus", fill="OTU" , facet_grid = type ~ pair) 
p$data$OTU    <- factor(p$data$OTU)
p$data$Genus  <- gsub(p$data$Genus, pattern="_etal", replacement="")
p_tetra_total <- p + 
  geom_bar(aes(fill = OTU, color = OTU), stat = "identity", position = "stack") +
  ylab("Read abundance (% total)") +
  xlab(label="") + 
  theme_bw() +
  theme(legend.position="none",
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8, vjust = 0.2),
        axis.text.x  = element_blank(),
        axis.text.y  = element_text(size = 6),
        axis.ticks.x = element_line(size = 0.3),
        axis.ticks.y = element_line(size = 0.3),
        strip.text   = element_text(size = 6),
        panel.grid.minor = element_blank(),
        strip.background = element_rect(linetype= "blank", fill = "white")
        )

p2 <- plot_bar(NOB.prop, "Genus", fill="OTU" , facet_grid = type ~ pair)
p2$data$OTU   <- factor(p2$data$OTU)
p2$data$Genus <- gsub(p2$data$Genus, pattern="_etal", replacement="")
p_tetra_rel <- p2 +
  geom_bar(aes(fill = OTU, color = OTU), stat = "identity", position = "stack") +
  ylab("Read abundance (% total)") +
  xlab(label="") + 
  theme_bw() +
  theme(legend.position="none",
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8, vjust = 0.2),
        axis.text.x  = element_text(size = 6,  hjust = 1, vjust = 1, 
                                    angle = 45),
        axis.text.y  = element_text(size = 6),
        axis.ticks.x = element_line(size = 0.3),
        axis.ticks.y = element_line(size = 0.3),
        strip.text   = element_text(size = 6),
        panel.grid.minor = element_blank(),
        strip.background = element_rect(linetype= "blank", fill = "white")
        )

pdf(file = "figs/Figure_S10.pdf")
grid.arrange(p_tetra_total, p_tetra_rel, nrow = 2)
dev.off()

```




